---
  title: "Untitled"
author: "Andrew W. Park"
date: '2023-04-07'
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's make a model!

$$
  \begin{aligned}
\frac{dS}{dt}&=ef A(S+I)-d S - \beta S Z\\
\frac{dI}{dt}&=\beta SZ-d_{i}I\\
\frac{dZ}{dt}&=d_{i}I\sigma_{max}(\frac{A}{A+h})-mZ -f(S+I)Z\\
\frac{dA}{dt}&=r_{A}A(1-\frac{A}{K_{A}})-f(S+I)A
\end{aligned}
$$
### What do each of these variables mean?   
S = susceptible Daphnia
I = infected Daphnia 
Z = metsch spores 
A = ank algae
e = conversion efficiency 
f = foraging rate
d = death rate of uninfected Daphnia
di = death rate of infected Daphnia
Beta = transmission parameter
sigma max = maximum spore yield
h = half-saturation constant of algae growth
m = spore loss rate 
rA = algal specific growth rate  
KA = algal carrying capacity


### Which of these things vary with temperature? 
$$
 \begin{align}
f(T,L) &= L^{\gamma}\cdot f_R \cdot e^{T_A([1/T_R]-[1/T])} \\
\beta_{adult}(T)&=\beta_R \cdot e^{T_A([1/T_R]-[1/T])} \\
u &= \frac{\beta_{adult}}{f_{adult}}\\
\beta_{pop} &= u \cdot f_{pop} \\
d(T) &= d_R \cdot e^{T_A([1/T_R]-[1/T])}\\
d_i(T) &= d_{iR} \cdot e^{T_A([1/T_R]-[1/T])}\\
r(T) &= r_R \cdot e^{T_A([1/T_R]-[1/T])}\\
b &= r+d\\
e &= \frac{b}{f \cdot A}\\
\sigma_{max}(T) &=\alpha_2 \cdot T^2+\alpha_1 \cdot T + \alpha_0\\
r_A(T) &= c e^{(A-\frac{E_a}{R}\cdot \frac{1}{T})}\\
K_A(T) &= c(\alpha_1 \cdot T+\alpha_0)\\
\end{align}
$$
Assumed to not be dependent on temperature: m and h
 
```{r, libload}
library(tidyverse)
library(magrittr)
library(deSolve)
```



```{r defineModel}

sirMod = function(t, y, params){
  S = y[1]
  I = y[2]
  R = y[3]
  
  
  with(as.list(params),{
    dS = lambda - mu*S - beta*S*I
    dI = beta*S*I - mu*I - gamma*I
    dR = gamma*I - mu*R 
    
    res=c(dS, dI, dR)
    list(res)
  }
  )
}

```

```{r defineParams}
times  = seq(0, 10, by=1/100)
params  = c(lambda=1000.0, mu=1.0, beta=0.02, gamma=12.0)
start = c(S = 500, I = 1, R = 0)
```




```{r calcR0}
R0 <- (params["beta"]*params["lambda"]/params["mu"])/(params["mu"]+params["gamma"])
```



```{r runModel}
out <- as.data.frame(ode(start, times, sirMod, params))
```

```{r plotDyn}
q <- out %>% pivot_longer(.,2:4,names_to="stateVar",values_to="size")
q %>% dplyr::filter(time<10) %>% ggplot(.,aes(x=time,y=size,col=stateVar))+geom_line()+theme_classic()
```

```{r loopParams}
myBetas <- seq(0.01,0.1,0.01)
myGammas <- seq(52.0,12.0,length.out=10) #from 1 week to one month

eq <- tibble(beta=numeric(0),gamma=numeric(0),eqPrev=numeric(0),R0=numeric(0))

for (i in 1:length(myBetas)){
  for (j in 1:length(myGammas)){
    params["beta"] <- myBetas[i]
    params["gamma"] <- myGammas[j]
    R0 <- (params["beta"]*params["lambda"]/params["mu"])/(params["mu"]+params["gamma"])
    R0 <- unname(R0)
    R0 <- as.numeric(format(R0,digits=2))
    out2 <- as.data.frame(ode(start, times, sirMod, params))
    eqPrev <- out2 %>% slice(n()) %>% mutate(p=I/(S+I+R)) %>% pull(p)
    eq %<>% add_case(beta=myBetas[i],gamma=myGammas[j],eqPrev=eqPrev,R0=R0)
  }
}

#plot
eq %>% ggplot(.,aes(x=beta,y=gamma))+geom_tile(color="black",aes(fill=eqPrev))+scale_fill_gradient(low="white",high="red")+geom_text(aes(label=R0),color="black",size=4)+theme_minimal()

```

