---
title: "Ceriodaphnia temperature effects"
author: "Katie Schroeder"
date: "8/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###read in data and packages  
```{r}
setwd("~/Documents/GitHub/2022-thermal-zoop-experiments")
lt_mortality<-read.csv("lifetable_mortality.csv")
lt<-read.csv("raw_lifetable.csv")
lt_lengths<-read.csv("lifetable_day_5_lengths.csv")

#packages
library(ggplot2)
library(tidyverse)
```

###life table data: effects of temperature on cerio  
```{r}
#filter lifetable mortality data to just cerio and look at lifespan at each temperature
ggplot(filter(lt_mortality,species=="cerio"),aes(x=temp_id,y=lifespan,color=as.factor(mean_temp)))+geom_point()+theme_bw()+ylab("Lifespan (days)")+xlab("Temperature (deg C)")+facet_wrap(.~as.factor(inf))
#0 = uninfected, 1 = infected, NA = inf not assessed

#Many cerio dapnia removed at end of experiment so lifespan is greater than those graphed. Find number removed for each temperature and mean lifespan
cerio_removed <- lt_mortality %>% filter(species=="cerio"&removed=="1") %>% group_by(temp_id,inf) %>% summarize(cerio_removed = sum(removed)) %>% ggplot(aes(x=temp_id,y=cerio_removed))+geom_point()+theme_bw()
cerio_removed
```
Both uninfected and infected Cerio live longer at cooler temperatures. Infection reduces lifespan at all temperatures  

###Spore yield and temperature
```{r}
#First look at spore yield
ggplot(filter(lt_mortality,species=="cerio"&inf=="1"),aes(x=temp_id,y=spore_RAW))+geom_boxplot()+theme_bw()

#Now look at body size
ggplot(filter(lt_mortality,species=="cerio"&inf=="1"),aes(x=temp_id,y=length_RAW))+geom_boxplot()+theme_bw()

#Look at spore yield and body size
ggplot(filter(lt_mortality,species=="cerio"&inf=="1"),aes(x=temp_id,y=spore_RAW/length_RAW))+geom_boxplot()+theme_bw()
```
Ceriodaphnia tend to be bigger at higher temperatures but spore yield is pretty consistent across temperatures. Spore yield/body size is slightly elevated at higher temperatures

```{r}
cerio_summary <- lt_mortality %>% group_by(temp_id,inf) %>% summarize(mean_lifespan = mean(lifespan,na.rm=T),
                                                                  lifespan_se = sd(lifespan)/sqrt(length(lifespan)))

```

