---
title: "Kunze paper data re-work"
author: "Katie Schroeder"
date: "2023-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data availability.   
Data were pulled from Kunze et al. (2022) *eLife* paper titled "Alternate patterns of temperature variation bring about very different disease outcomes at different mean temperatures."    

All data are available on GitHub page. User: charlyknz in HostParasite repository.  

## Purpose of this analysis.  
Revisit the data collected in this experiment and determine if any of the rates can be calculated via rate summation/nonlinear averaging.

## Step 1: load in data and necessary libraries.  
```{r libraries and data loading in}
library(tidyverse)
library(magrittr)
library(ggpubr) #for ggerrorplot

spore <- read_csv("Kunze et al 2022/SporesNoMaleNA.csv")
spore %<>% filter(treatment!="PULSE") #remove the pulse treatments from the dataframe. Only interested in constant and fluctuating temperatures
spore %<>% filter(is.na(comment)) #removes commented samples - many killed by pipette, otherwise not reliable 

```

## Spore data frame characteristics:  
treatment is either fluctuating (FLU) or constant (CS).  
exposed is either U (uninfected/unexposed) or I (infected/exposed).  
all fluctuations were +/- 3 deg C every 12 hours.  
temperature info: have temperature, mean temperature, real mean temperature (from HOBO)  
death day: format %d/%m/%Y.  
infect: 0 or 1 on whether Daphnia were infected.  
no_spore: number of spores inside the host upon death - exclude before day 11, only include last day of experiment in paper.  

## Step 2: look at growth rate.  
```{r}
#to get growth rate, first need to calculate lifespan (assume birthday is 4/09)
spore %<>% mutate(birthday = "09/04/2019",
                 lifespan = as.double(as.Date(death,'%d/%m/%Y')-as.Date(birthday,'%d/%m/%Y')),
                 growth_rate = size/lifespan)

#graph out growth rate at each temperature, separating fluctuating and constant treatments
#filter out those younger than 26, when measurements began
spore %>% filter(lifespan >=25) %>% ggplot(aes(x=meantemp,y=growth_rate,color=treatment,group=treatment)) + geom_point() + theme_bw() + geom_line(data=preds_beta,aes(x=temp,y=.fitted),color='skyblue')

#summarize w/ error bars
ggerrorplot(spore,x="meantemp",y="growth_rate",desc_stat="mean_se",color="treatment",palette="Paired",facet.by="infect")
```
## Fit a beta function and see if it predicts growth rate.  
```{r}
library(rTPC)
library(nls.multstart)
library(broom)

#separate into uninfected and infected for the constant
uninf <- spore %>% filter(infect==0)
inf <- spore %>% filter(infect==1)

#filter out samples without a growth rate and only look at constant uninfected samples
d <- uninf %>% filter(treatment=="CS"&growth_rate>=0) %>% mutate(temp = meantemp, rate=growth_rate)

betafit<- nls_multstart(rate~beta_2012(temp = temp, a, b, c, d, e),
              data = d,
              iter = c(6,6,6,6,6),
              start_lower = get_start_vals(d$temp, d$rate, model_name = 'beta_2012') - 10,
              start_upper = get_start_vals(d$temp, d$rate, model_name = 'beta_2012') + 10,
              lower = get_lower_lims(d$temp, d$rate, model_name = 'beta_2012'),
              upper = get_upper_lims(d$temp, d$rate, model_name = 'beta_2012'),
              supp_errors = 'Y',
              convergence_count = FALSE)

new_data_beta <- data.frame(temp = seq(min(d$meantemp), max(d$meantemp), 0.1))
preds_beta <- augment(betafit, newdata = new_data_beta)

#graph
ggplot(d,aes(meantemp,growth_rate)) +
  geom_point() +
  geom_line(data=preds_beta,aes(x=temp,y=.fitted),color='skyblue')+
  theme_classic() 

#manipulate the beta functions predicted values to get a dataframe that can be used to fit Jensen-esque line
beta_preds <- preds_beta %>% filter(temp=="10"|temp=="13"|temp=="16"|temp=="19"|temp=="22"|temp=="25"|temp=="28")

write.csv(beta_preds,"beta_preds.csv")

#get summary info on growth rate so that we can compare fitted to estimates
spore_sum <- spore %>% filter(growth_rate>=0) %>% group_by(meantemp,treatment,infect) %>% 
  summarize(mean_growthrate = mean(growth_rate),
            count = n(),
            se_growthrate = sd(growth_rate)/sqrt(count))

spore_sum_sum <- spore_sum %>% filter(treatment == "FLU" & infect ==0)

#build new df
estimates <- data.frame(temp1 = c(10,13,16,19,22),
                        fitted1 = c(beta_preds$.fitted[1],beta_preds$.fitted[2],beta_preds$.fitted[3],beta_preds$.fitted[4],beta_preds$.fitted[5]),
                        temp2 = c(16,19,22,25,28),
                        fitted2 = c(beta_preds$.fitted[3],beta_preds$.fitted[4],beta_preds$.fitted[5],beta_preds$.fitted[6],beta_preds$.fitted[7]),
                        fluct = c(13,16,19,22,25),
                        actual = spore_sum_sum$mean_growthrate,
                        actual_se = spore_sum_sum$se_growthrate,
                        actual_count = spore_sum_sum$count)

estimates %<>% mutate(midpoint = (fitted1+fitted2)/2)

estimates %>% ggplot(aes(x=fluct,y=actual)) + 
  geom_point() + 
  geom_point(aes(x=fluct,y=midpoint),color="blue") + 
  geom_errorbar(aes(ymin=actual-actual_se,ymax=actual+actual_se),width=0.1) + 
  geom_line(data=preds_beta,aes(x=temp,y=.fitted),color='skyblue')+
  theme_classic() 


```
What about when things are infected?
```{r}
#filter out samples without a growth rate and only look at constant infected samples
d <- inf %>% filter(treatment=="CS"&growth_rate>=0) %>% mutate(temp = meantemp, rate=growth_rate)

betafit<- nls_multstart(rate~beta_2012(temp = temp, a, b, c, d, e),
              data = d,
              iter = c(6,6,6,6,6),
              start_lower = get_start_vals(d$temp, d$rate, model_name = 'beta_2012') - 10,
              start_upper = get_start_vals(d$temp, d$rate, model_name = 'beta_2012') + 10,
              lower = get_lower_lims(d$temp, d$rate, model_name = 'beta_2012'),
              upper = get_upper_lims(d$temp, d$rate, model_name = 'beta_2012'),
              supp_errors = 'Y',
              convergence_count = FALSE)

new_data_beta <- data.frame(temp = seq(min(d$meantemp), max(d$meantemp), 0.1))
preds_beta <- augment(betafit, newdata = new_data_beta)

#graph
ggplot(d,aes(meantemp,growth_rate)) +
  geom_point() +
  geom_line(data=preds_beta,aes(x=temp,y=.fitted),color='skyblue')+
  theme_classic() 

#manipulate the beta functions predicted values to get a dataframe that can be used to fit Jensen-esque line
beta_preds <- preds_beta %>% filter(temp=="10"|temp=="13"|temp=="16"|temp=="19"|temp=="22"|temp=="25"|temp=="28")

write.csv(beta_preds,"beta_preds.csv")

#get summary info on growth rate so that we can compare fitted to estimates
spore_sum <- spore %>% filter(growth_rate>=0) %>% group_by(meantemp,treatment,infect) %>% 
  summarize(mean_growthrate = mean(growth_rate),
            count = n(),
            se_growthrate = sd(growth_rate)/sqrt(count))

spore_sum_sum <- spore_sum %>% filter(treatment == "FLU" & infect ==1)

#build new df
estimates <- data.frame(temp1 = c(10,13,16,19,22),
                        fitted1 = c(beta_preds$.fitted[1],beta_preds$.fitted[2],beta_preds$.fitted[3],beta_preds$.fitted[4],beta_preds$.fitted[5]),
                        temp2 = c(16,19,22,25,28),
                        fitted2 = c(beta_preds$.fitted[3],beta_preds$.fitted[4],beta_preds$.fitted[5],beta_preds$.fitted[6],beta_preds$.fitted[7]),
                        fluct = c(13,16,19,22,25),
                        actual = spore_sum_sum$mean_growthrate,
                        actual_se = spore_sum_sum$se_growthrate,
                        actual_count = spore_sum_sum$count)

estimates %<>% mutate(midpoint = (fitted1+fitted2)/2)

estimates %>% ggplot(aes(x=fluct,y=actual)) + 
  geom_point() + 
  geom_point(aes(x=fluct,y=midpoint),color="blue") + 
  geom_errorbar(aes(ymin=actual-actual_se,ymax=actual+actual_se),width=0.1) + 
  geom_line(data=preds_beta,aes(x=temp,y=.fitted),color='skyblue')+
  theme_classic() 
```





```{r}

sloperify <- function(i, j){
  m <- (beta_preds$.fitted[j] - beta_preds$.fitted[i])/(beta_preds$temp[j] - beta_preds$temp[i])
  b <- -m * beta_preds$temp[i] + beta_preds$.fitted[i]
  mid <- (beta_preds$temp[j] - beta_preds$temp[i])/2 + beta_preds$temp[i]
  y <- m * mid + b
  return(y)
}

sloperify <- function(i, j, k, l){
  m <- (j - i)/(k - l)
  b <- -m * l + i
  mid <- (beta_preds$temp[j] - beta_preds$temp[i])/2 + beta_preds$temp[i]
  y <- m * mid + b
  return(y)
}

```

