---
title: "2022 Tank Experiment REU + a month"
author: "Katie Schroeder"
date: "8/2/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
General notes:
##trapZ - integrated area under the curve. As time series
##Coefficient of variation to look at how much they cycle - standard deviation/mean
##Jessica Hite - dissertation chapter - synchrony and size structure
##Infection prev vs density of inf hosts. Diff metric of disease may matter. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Experiment Overview

###Boring set up stuff
Packages
```{r, message = FALSE,warning=FALSE}
library(tidyverse)
library(magrittr)
library(ggnewscale)
library(lubridate)
```

Reading in data
Most recent copies of data live in the drive - links in the experiment protocol. YSI data lives in Dropbox.
```{r, message = F}
setwd("/Users/katieschroeder/Documents/GitHub/2022-thermal-zoop-experiments/raw-data/tank-expt")
meta <- read.csv("tank_metadata.csv")
counts <- read.csv("tank_counts.csv")
ysi <- read_csv("tank_ysi.csv")
```

Fixing YSI data to look normal
```{r}
colnames(ysi)[5:24] <- c("temp_f","mmHg","do_per","do_conc","spc","c-us","nlfc",
                         "ohm","tds","sal","density","density_2","pH","pH_mv",
                         "phyco_rfu","phyco_ug","chlor_rfu","chlor_ug","DEP","vpos")


ysi %<>% mutate(bucket = as.factor(as.numeric(gsub("Tank ", "", Site))))
ysi %<>% mutate(temp_c = (temp_f-32)*(5/9))
ysi$Date <- mdy(ysi$Date)
```

```{r}
meta$bucket <- as.factor(meta$bucket)
counts$bucket <- as.factor(counts$bucket)
```

##Zooplankton Data
Set up total, daphnia and cerio prevalence calculations:  
Total = sum of all zooplankton (cerio+daphnia)  
Daphnia = sum of all Daphnia (inf + uninf, juv + adult, males included)  
Cerio = sum of all Cerio (inf + uninf)  
Daphnia_inf = inf juveniles and adults  
Daphnia_sus = uninf juveniles and adults  
Daphnia_prev = inf Daphnia / total number of Daphnia  
Cerio_prev = inf Cerio / total number of Cerio  
D_ratio = number of Daphnia / total Zooplankton (% Daphnia)  
```{r}
data<-left_join(counts,meta, by="bucket")

data$community <- as.factor(data$community)
data$sample_date <- mdy(data$sample_date)
data$zoop_date <- mdy(data$zoop_date)
data$spore_date <- mdy(data$spore_date)
data %<>% mutate(total = rowSums(across(juvenile_daphnia_uninf:cerio_inf),na.rm=T), 
                 daphnia = rowSums(across(juvenile_daphnia_uninf:male_daphnia_inf),na.rm=T),
                 cerio = rowSums(across(cerio_uninf:cerio_inf)),
                 treatment = paste(community, temp, sep="_"),
                 daphnia_inf = rowSums(across(contains("daphnia_inf")),na.rm=T),
                 daphnia_sus = rowSums(across(contains("daphnia_uninf")),na.rm=T),
                 daphnia_prev = daphnia_inf/daphnia,
                 daphnia_inf_status = ifelse(daphnia_prev>0,"inf","uninf"),
                 cerio_prev = cerio_inf/cerio,
                 cerio_inf_status = ifelse(cerio_prev>0,"inf","uninf"),
                 d_ratio = daphnia/total,
                 juveniles = juvenile_daphnia_inf+juvenile_daphnia_uninf,
                 j_ratio = juveniles/daphnia)

#total zoop for ESA
total_summary <- data %>% 
  group_by(temp,sample_date,community) %>% 
  dplyr::summarise(total = mean(total,na.rm=TRUE),
            count = n(),
              se = sd(total)/sqrt(count))

data %>% filter(community=="DM"|community=="D") %>%
  filter(bucket!="30"&bucket!="26"&bucket!="23"&bucket!="22") %>%
  ggplot(aes(x=sample_date,y=total))+ 
  geom_point(color="lightgrey") +
  geom_line(aes(group = bucket),color="lightgrey") + 
  geom_point(data=filter(total_summary,community=="DM"|community=="D"),color="black") + 
  geom_line(data=filter(total_summary,community=="DM"|community=="D"),color="black")+
  scale_color_discrete() + 
  facet_grid(rows=vars(community),cols=vars(temp)) +
  theme_classic() + 
  ylab("Population Size (# of zoops)") +
  theme(text = element_text(size = 20),legend.position="none")+
  xlab("Sampling Date")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

data %>% filter(community=="D") %>%
  filter(bucket!="30"&bucket!="26"&bucket!="23"&bucket!="22") %>%
  ggplot(aes(x=sample_date,y=total))+ 
  geom_point(color="lightgrey") +
  geom_line(aes(group = bucket),color="lightgrey") + 
  geom_point(data=filter(total_summary,community=="D"),color="black") + 
  geom_line(data=filter(total_summary,community=="D"),color="black")+
  scale_color_discrete() + 
  facet_wrap(~temp,nrow=1) +
  theme_classic() + 
  ylab("Population Size (# of zoops)") +
  theme(text = element_text(size = 20),legend.position="none")+
  xlab("Sampling Date")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#Time to try and get days between sample date and the start date
data %<>% mutate(days_since_hosts_added = as.Date(sample_date)-as.Date(zoop_date),
                 days_since_spores_added = as.Date(sample_date)-as.Date(spore_date))
#Set up another version of the data frame that has infected and susceptible Daphnia counts in one column for easier SI graphing
data_long <- pivot_longer(data,cols = c("daphnia_inf","daphnia_sus"),names_to="daphnia_SI_status",values_to="daphnia_SI_count")
```

```{r,eval=FALSE,include=FALSE}
#This chunk can be used to make the super messy graphs that show all buckets separated by community
ggplot(data,aes(x=sample_date,y=total, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + theme(legend.position = "none") + facet_wrap(vars(community))+theme_bw()
ggplot(data,aes(x=sample_date,y=daphnia, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() +  facet_wrap(vars(community))+theme_bw() 
ggplot(data,aes(x=sample_date,y=cerio_uninf, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_wrap(vars(community))+theme_bw() 
```

###Totals (All zoops, Daphnia, Cerio) faceted by temp treatment and community composition  
How do densities of zooplankton vary between temperature and different community compositions?  
```{r}
ggplot(data,aes(x=sample_date,y=total, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + theme(legend.position = "none") + facet_grid(rows= vars(temp), cols = vars(community)) + theme_bw() + ylab("Total zoop density") + ggtitle("Total Zoop")
ggplot(data,aes(x=sample_date,y=daphnia, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Total Daphnia density") + ggtitle("Daphnia")
ggplot(filter(data,community=="CM"|community=="DCM"),aes(x=sample_date,y=cerio_uninf, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Total Cerio density") + ggtitle("Cerio")
```

####Now doing it all again on a log scale  
```{r,message=FALSE,warning=FALSE}
ggplot(data,aes(x=days_since_hosts_added,y=total))+ 
  geom_point() + 
  geom_line(aes(group = bucket)) + 
  scale_color_discrete() + 
  theme(legend.position = "none") + 
  facet_grid(rows= vars(temp), cols = vars(community)) + 
  theme_bw() + ylab("Total zoop density (log scale)") + 
  ggtitle("Total Zoop") + scale_y_continuous(trans='log10')

data %>%
  filter(community!="CM") %>%
  ggplot(aes(x=days_since_hosts_added,y=daphnia))+ 
  geom_line(aes(group = bucket)) + 
  geom_point(size=0.5)+
  scale_color_discrete() + 
  facet_grid(rows= vars(temp), cols = vars(community)) +
  theme_bw() + ylab("Total Daphnia density (log scale)") + 
  ggtitle("Daphnia") + scale_y_continuous(trans='log10')

ggplot(filter(data,community=="CM"|community=="DCM"),aes(x=sample_date,y=cerio_uninf, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Total Cerio density (log scale)") + ggtitle("Cerio") + scale_y_continuous(trans='log10')
```
####Which buckets have bottomed out?
```{r}
no_daphnia<- data %>% filter(community!="CM"&daphnia=="0")
unique(no_daphnia$bucket)

no_cerio <- data %>% filter(community=="CM"&cerio=="0")
unique(no_cerio$bucket)

#Make summary statistics for density after July 15
density_summary_omit <- data %>% filter(bucket!="22"|bucket!="23"|bucket!="26"|bucket!="30"|bucket!="43") %>% group_by(treatment) %>% summarize(mean_cerio_o = mean(cerio),
                                                                se_cerio = sqrt(var(cerio,na.rm=T)/n()),
                                                                mean_daphnia = mean(daphnia),
                                                                se_daphnia = sqrt(var(daphnia,na.rm=T)/n()))
```


####Which CM buckets are contaminated with Daphnia?
```{r}
contaminated_CM <- data %>% filter(community=="CM"&daphnia>2)
unique(contaminated_CM$bucket)
```
CM buckets for each temperature:
15C: 33-36
20C: 37-40
25C: 41-44
Var: 45-48

####Taking total density data and incorporating if any Daphnia or Ceriodaphnia were infected during that sampling
```{r}
ggplot(filter(data,community!="CM"),aes(x=sample_date,y=daphnia, color=daphnia_inf_status))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Total Daphnia density") + ggtitle("Daphnia")

ggplot(filter(data,community=="CM"|community=="DCM"),aes(x=sample_date,y=cerio_uninf, color=cerio_inf_status))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Total Cerio density") + ggtitle("Cerio")

#Can do color as continuous prevalence but it's harder to see which buckets had infections
```

###Infection data
####Infected individuals
```{r}
ggplot(data,aes(x=sample_date,y=daphnia_inf, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Daphnia Inf Individuals (log scale)") + ggtitle("Daphnia I") + scale_y_continuous(trans='log10')

#for ESA:combining DM, DCM
data %<>% mutate(temp = if_else(temp=="var","Fluctuating",temp))

data %>% filter(bucket == "20" | bucket == "53" | bucket == "44" | bucket == "63") %>%
  ggplot(aes(x=sample_date,y=daphnia_inf))+ 
  geom_point(size=3) + 
  geom_line(aes(group = bucket),size=1.5) + 
  scale_color_discrete() + 
  facet_wrap(.~temp) +
  theme_classic() + 
  ylab(expression('Number of Infected'~italic(Daphnia))) + 
  xlab("Sample date")+
  theme(text = element_text(size = 20),legend.position="none")
ggsave("inf_over_time.png",dpi=500,height=7.5,width=13,units="in")

data %>% filter(bucket == "20" | bucket == "53" | bucket == "44" | bucket == "63") %>%
  ggplot(aes(x=sample_date,y=daphnia_inf))+ 
  geom_point(size=3) + 
  geom_line(aes(group = bucket),size=1.5) + 
  scale_color_discrete() + 
  facet_wrap(.~temp) +
  theme_classic() + 
  ylab(expression('Number of Infected'~italic(Daphnia))) + 
  xlab("Sample date")+
  theme(text = element_text(size = 20),legend.position="none") +
  geom_area(fill="seagreen3")
ggsave("inf_over_time_shade.png",dpi=500,height=7.5,width=13,units="in")

ggplot(filter(data,community=="CM"|community=="DCM"),aes(x=sample_date,y=cerio_inf, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Cerio Inf Individuals (log scale)") + ggtitle("Cerio I") + scale_y_continuous(trans='log10')

#Cerio density vs cerio inf
ggplot(filter(data,community=="CM"&daphnia<=3),aes(x=cerio,y=cerio_inf,color=temp))+ geom_point() + scale_color_discrete() + facet_wrap(.~temp) +theme_bw() + ylab("Cerio Inf Individuals") + ggtitle("Cerio Inf/Density")
```

####Prevalence
```{r,message=FALSE,warning=FALSE}
data %>% filter(community=="DCM"|community=="DM") %>%
ggplot(aes(x=sample_date,y=daphnia_prev))+ 
  geom_point() + 
  geom_line(aes(group = bucket)) + 
  scale_color_discrete() + 
  facet_wrap(~temp,nrow=1) +
  theme_classic() + 
  ylab("Infection Prevalence") +
  xlab("Sampling Date")

prev_summary <- data %>% filter(community=="DCM"|community=="DM") %>%
  group_by(temp,sample_date) %>% dplyr::summarize(daphnia_prev = mean(daphnia_prev,na.rm=TRUE),
                                           se = sd(daphnia_prev)/sqrt(n()))

prev_summary %>% 
 ggplot(aes(x=temp,y=daphnia_prev)) +
  geom_boxplot(color="lightgrey") +
  geom_point(size=2,color="black",alpha=2) +
  theme_classic()


ggplot(data=filter(data,community=="DCM"|community=="DM"),aes(x=sample_date,y=daphnia_prev))+ 
  geom_point(color="lightgrey") +
  geom_line(data=data,aes(group = bucket),color="lightgrey") + 
  geom_point(data=prev_summary,color="black") + 
  geom_line(data=prev_summary,color="black")+
  scale_color_discrete() + 
  facet_wrap(~temp,nrow=1) +
  theme_classic() + 
  ylab("Infection Prevalence") +
  theme(text = element_text(size = 20),legend.position="none")+
  xlab("Sampling Date")

ggplot(filter(data,community=="CM"|community=="DCM"),aes(x=sample_date,y=cerio_prev, color=temp))+ geom_point() + geom_line(aes(group = bucket)) + scale_color_discrete() + facet_grid(rows= vars(temp), cols = vars(community)) +theme_bw() + ylab("Cerio Inf Prev (log scale)") + ggtitle("Cerio Prev")
```

###Summary data
```{r,include=F}
#try filtering out 26 and 30 for 25 and var
summary <- data %>% group_by(treatment, sample_date) %>% 
  dplyr::summarize(mean_juvenile = mean(juvenile_daphnia_uninf),
            se_juvenile_daphnia_uninf=sqrt(var(juvenile_daphnia_uninf,na.rm=T)/n()),
            mean_adult = mean(adult_daphnia_uninf),
            se_adult = sqrt(var(adult_daphnia_uninf,na.rm=T)/n()),
            mean_cerio = mean(cerio_uninf),
            se_cerio = sqrt(var(cerio_uninf,na.rm=T)/n()),
            mean_daphnia = mean(daphnia),
            se_daphnia = sqrt(var(daphnia,na.rm=T)/n()),
            mean_total = mean(total),
            se_total = sqrt(var(total,na.rm=T)/n()),
            mean_daphnia_inf = mean(daphnia_inf),
            se_daphnia_inf = sqrt(var(daphnia_inf,na.rm=T)/n()),
            mean_cerio_inf = mean(cerio_inf),
            se_cerio_inf = sqrt(var(cerio_inf,na.rm=T)/n()),
            mean_daphnia_prev = mean(daphnia_prev),
            se_daphnia_prev = sqrt(var(daphnia_prev,na.rm=T)/n()),
            mean_cerio_prev = mean(cerio_prev),
            se_cerio_prev = sqrt(var(cerio_prev,na.rm=T)/n()),
            mean_d_ratio = mean(d_ratio),
            se_d_ratio = sqrt(var(d_ratio,na.rm=T)/n()),
            mean_j_ratio = mean(j_ratio),
            se_j_ratio = sqrt(var(j_ratio,na.rm=T)/n()),
            temp = temp,
            community = community)
```

####Summary graphs by sampling date for temperature trends in density and infection
Spores were introduced 7/11/2022, first infections observed 7/18/2022 

```{r,message=F,warning=FALSE,eval=FALSE,include=FALSE}
ggplot(summary,aes(x=temp,y=mean_total,color=community))+geom_point()+geom_pointrange(aes(ymin=mean_total-se_total,ymax=mean_total+se_total))+geom_line(aes(group=community))+facet_wrap(.~sample_date,nrow=2)+theme_bw()+ scale_y_continuous(trans='log10')+ylab("Total Zoop Density (log scale)")

ggplot(filter(summary,community!="CM"),aes(x=temp,y=mean_daphnia,color=community))+geom_point()+geom_pointrange(aes(ymin=mean_daphnia-se_daphnia,ymax=mean_daphnia+se_daphnia))+geom_line(aes(group=community))+facet_wrap(.~sample_date,nrow=2)+theme_bw()+ scale_y_continuous(trans='log10')+ylab("Total Daphnia Density (log scale)")

##By sample date
ggplot(filter(summary,community!="CM"),aes(x=sample_date,y=mean_daphnia,color=temp))+geom_point()+geom_pointrange(aes(ymin=mean_daphnia-se_daphnia,ymax=mean_daphnia+se_daphnia))+geom_line(aes(group=temp))+facet_wrap(.~community,nrow=1)+theme_bw()+ scale_y_continuous(trans='log10')+ylab("Total Daphnia Density (log scale)")

ggplot(filter(summary,community=="CM"|community=="DCM"),aes(x=temp,y=mean_cerio,color=community))+geom_point()+geom_pointrange(aes(ymin=mean_cerio-se_cerio,ymax=mean_cerio+se_cerio))+geom_line(aes(group=community))+facet_wrap(.~sample_date,nrow=2)+theme_bw()+ scale_y_continuous(trans='log10')+ylab("Total Cerio Density (log scale)")
```

####Infection summary graphs   
```{r,message=F,warning=FALSE}

ggplot(filter(summary,community!="CM"),aes(x=sample_date,y=mean_daphnia_prev,color=temp))+geom_point()+geom_pointrange(aes(ymin=mean_daphnia_prev-se_daphnia_prev,ymax=mean_daphnia_prev+se_daphnia_prev))+geom_line(aes(group=temp))+facet_wrap(.~community,nrow=1)+theme_bw()+ylab("Daphnia Inf Prev")

ggplot(filter(summary,community=="CM"|community=="DCM"),aes(x=sample_date,y=mean_cerio_prev,color=temp))+geom_point()+geom_pointrange(aes(ymin=mean_cerio_prev-se_cerio_prev,ymax=mean_cerio_prev+se_cerio_prev))+geom_line(aes(group=temp))+facet_wrap(.~community,nrow=1)+theme_bw()+ylab("Cerio Inf Prev")
```
###Epidemic Size - measuring area under the epidemic curve
```{r}
#First visualize epidemics in each temperature and community treatment again - ignore CM for now
data %>% filter(days_since_spores_added>=0&community!="CM") %>% ggplot(.,aes(x=days_since_spores_added,y=daphnia_prev,color=temp))+
  geom_point()+
  geom_line(aes(group = bucket)) + 
  facet_grid(rows= vars(temp), cols = vars(community))+
  theme_bw()

#To use trapz
library(pracma)

#Try it as a for loop ):
buckets <- unique(data$bucket)
epidemic_size_df <- data.frame(bucket <- buckets)

for (i in 1:length(buckets)){
  sample_data <- data[data$bucket==buckets[i],]
  #relevant extra info
  epidemic_size_df$temp[i] <- as.character(sample_data$temp[1])
  epidemic_size_df$community[i] <- as.character(sample_data$community[1])
  epidemic_size_df$treatment[i] <- as.character(sample_data$treatment[1])
  #now try to actually integrate
  epidemic_size_df$epidemic_size[i] <- trapz(as.numeric(sample_data$days_since_spores_added),matrix(as.numeric(sample_data$daphnia_prev)))
  epidemic_size_df$epidemic_size_ind[i] <- trapz(as.numeric(sample_data$days_since_spores_added),matrix(as.numeric(sample_data$daphnia_inf)))
  #add cerio in
  epidemic_size_df$epidemic_size_c[i] <- trapz(as.numeric(sample_data$days_since_spores_added),matrix(as.numeric(sample_data$cerio_prev)))
  epidemic_size_df$epidemic_size_ind_c[i] <- trapz(as.numeric(sample_data$days_since_spores_added),matrix(as.numeric(sample_data$cerio_inf)))
}

#making the epidemic_size_df nicer to work with
epidemic_size_df %<>% rename(bucket = bucket....buckets) 

#ESA figure
epidemic_size_df %>% filter(community=="DM"|community=="DCM") %>% filter(epidemic_size_ind>10) %>% ggplot(aes(x=temp,y=epidemic_size_ind)) + 
  geom_boxplot() + 
  labs(x="Temperature Treatment",y="Total Epidemic Size (Number Infected)") + 
  geom_jitter(shape = 16, position =position_jitter(0.1),alpha=0.3)+theme_classic()

#raw data
ggplot(filter(epidemic_size_df,community!="CM"),aes(x=temp,y=epidemic_size,color=temp))+
  geom_point()+
  facet_wrap(.~community)+
  theme_bw()
ggplot(filter(epidemic_size_df,community!="CM"),aes(x=temp,y=epidemic_size_ind,color=temp))+geom_point()+facet_wrap(.~community)+theme_bw()

#summarized epidemic data
summary_ep <- filter(epidemic_size_df,bucket!="26"&bucket!="30"&bucket!="22"&bucket!="23"&bucket!="43") %>% group_by(treatment) %>% summarize(mean_epidemic_size = mean(epidemic_size,na.rm=T),
                                                                se_epidemic_size=sqrt(var(epidemic_size)/n()),
                                                                mean_epidemic_size_ind = mean(epidemic_size_ind,na.rm=T),
                                                                se_epidemic_size_ind=sqrt(var(epidemic_size_ind)/n()),
                                                                mean_epidemic_size_c = mean(epidemic_size_c,na.rm=T),
                                                                se_epidemic_size_c=sqrt(var(epidemic_size_c)/n()),
                                                                mean_epidemic_size_ind_c = mean(epidemic_size_ind_c,na.rm=T),
                                                                se_epidemic_size_ind_c=sqrt(var(epidemic_size_ind_c)/n()),
                                                                temp = unique(temp),
                                                                community = unique(community))
ggplot(filter(summary_ep,community!="CM"),aes(x=temp,y=mean_epidemic_size,color=temp))+geom_point()+facet_wrap(.~community)+theme_bw()+geom_pointrange(aes(ymin=mean_epidemic_size-se_epidemic_size,ymax=mean_epidemic_size+se_epidemic_size)) + ylab("daphnia epidemic size (prevalence days)")

ggplot(filter(summary_ep,community!="CM"),aes(x=temp,y=mean_epidemic_size_ind,color=temp))+geom_point()+facet_wrap(.~community)+theme_bw()+geom_pointrange(aes(ymin=mean_epidemic_size_ind-se_epidemic_size_ind,ymax=mean_epidemic_size_ind+se_epidemic_size_ind)) + ylab("epidemic size (total inf daphnia)")

ggplot(filter(summary_ep,community=="CM"|community=="DCM"),aes(x=temp,y=mean_epidemic_size_c,color=temp))+geom_point()+facet_wrap(.~community)+theme_bw()+geom_pointrange(aes(ymin=mean_epidemic_size_c-se_epidemic_size_c,ymax=mean_epidemic_size_c+se_epidemic_size_c)) + ylab("cerio epidemic size (prevalence days)")

ggplot(filter(summary_ep,community=="CM"|community=="DCM"),aes(x=temp,y=mean_epidemic_size_ind_c,color=temp))+geom_point()+facet_wrap(.~community)+theme_bw()+geom_pointrange(aes(ymin=mean_epidemic_size_ind_c-se_epidemic_size_ind_c,ymax=mean_epidemic_size_ind_c+se_epidemic_size_ind_c)) + ylab("total inf cerio")

##FOR GRFP
ggplot(filter(summary_ep,community=="DCM"),aes(temp,y=mean_epidemic_size))+geom_point()+theme_bw()+theme(legend.position = "none")+geom_pointrange(aes(ymin=mean_epidemic_size-se_epidemic_size,ymax=mean_epidemic_size+se_epidemic_size)) + ylab("Epidemic Size (prevalence*days)")+labs(x=expression(Temperature~(degree*C)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
####How does density at the beginning of the epidemic affect epidemic size? Look at densities before first infections were identified and then compare to epidemic size
```{r}
#first get data frame with density of daphnia and cerio before first infections (7/14)
density_epidemic <- data %>% filter(sample_date=="2022-07-14") %>% select(sample_date,bucket,daphnia,cerio,treatment) %>% left_join(epidemic_size_df,by="bucket")

ggplot(filter(density_epidemic,community!="CM"&community!="D"),aes(x=daphnia,y=epidemic_size,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)

ggplot(filter(density_epidemic,community!="CM"&community!="D"),aes(x=daphnia,y=epidemic_size_ind,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)

ggplot(filter(density_epidemic,community=="CM"|community=="DCM"),aes(x=cerio,y=epidemic_size_c,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)

ggplot(filter(density_epidemic,community=="CM"|community=="DCM"),aes(x=cerio,y=epidemic_size_ind_c,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)

#remove buckets that bottomed out
density_epidemic <- filter(density_epidemic,bucket!="22",bucket!="23"|bucket!="26"|bucket!="30"|bucket!="43")
ggplot(filter(density_epidemic,community!="CM"&community!="D"),aes(x=daphnia,y=epidemic_size,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)

ggplot(filter(density_epidemic,community!="CM"&community!="D"),aes(x=daphnia,y=epidemic_size_ind,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)

ggplot(filter(density_epidemic,community=="CM"|community=="DCM"),aes(x=cerio,y=epidemic_size_c,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)

ggplot(filter(density_epidemic,community=="CM"|community=="DCM"),aes(x=cerio,y=epidemic_size_ind_c,color=temp))+
  geom_point()+
  theme_bw()+
  facet_wrap(.~community)
```


###Daphnia and Cerio ratios and 20C vs Var infection for Jenavier's project
####Daphnia:cerio ratio for all treatments
```{r}
data %>% filter(community=="DCM") %>% ggplot(., aes(x=sample_date,y=d_ratio, color = temp)) + geom_point() +geom_line(aes(group = bucket)) + facet_wrap(vars(temp))+theme_bw()
```

For Jenavier, how do D to C ratios vary in var vs constant 20C temp buckets?
```{r}
#D to C ratio graph for all temperature DCM buckets
ggplot(filter(summary,community=="DCM"),aes(x=sample_date,y=mean_d_ratio,color=temp))+
  geom_point()+
  geom_pointrange(aes(ymin=mean_d_ratio-se_d_ratio,ymax=mean_d_ratio+se_d_ratio))+
  theme_classic()+
  facet_wrap(~temp)+
  xlab("Treatment")+
  ylab("Mean ratio of Daphnia to Ceriodaphnia")

ggplot(filter(data,community=="DCM"),aes(x=sample_date,y=d_ratio,color=temp,group=bucket))+
  geom_point()+
  geom_line() +
  theme_classic()+
  facet_wrap(~temp)+
  xlab("Treatment")+
  ylab("Ratio of Daphnia to Ceriodaphnia")

ggplot(filter(summary,community=="DCM"),aes(x=sample_date,y=mean_d_ratio,color=temp))+
  geom_point()+
  geom_pointrange(aes(ymin=mean_d_ratio-se_d_ratio,ymax=mean_d_ratio+se_d_ratio))+
  theme_classic()+
  facet_wrap(~temp)+
  xlab("Treatment")+
  ylab("Mean ratio of Daphnia to Ceriodaphnia")

#D to C ratio graph for just the 20 and var DCM buckets
#ggplot(filter(summary,community=="DCM"&(temp=="20"|temp=="var")),aes(x=temp,y=mean_d_ratio,color=temp))+geom_pointrange(aes(ymin=mean_d_ratio-se_d_ratio,ymax=mean_d_ratio+se_d_ratio))+facet_wrap(.~sample_date,nrow=1)+theme_bw()+xlab("Treatment")+ylab("Mean ratio of Daphnia to Ceriodaphnia")
ggplot(filter(summary,community=="DCM"&(temp=="20"|temp=="var")),aes(x=sample_date,y=mean_d_ratio,color=temp))+geom_pointrange(aes(ymin=mean_d_ratio-se_d_ratio,ymax=mean_d_ratio+se_d_ratio))+geom_line(aes(group=temp))+theme_bw()+xlab("Treatment")+ylab("Mean ratio of Daphnia to Ceriodaphnia")

#Infection Prevalence data
ggplot(filter(summary,community=="DCM"&(temp=="20"|temp=="var")),aes(x=sample_date,y=mean_daphnia_prev,color=temp))+geom_point()+geom_errorbar(aes(ymin=mean_daphnia_prev-se_daphnia_prev,ymax=mean_daphnia_prev+se_daphnia_prev),width=0.2)+geom_line(aes(group=temp))+theme_bw()+xlab("Treatment")+ylab("Mean inf prev")
```

###Epidemic size - integrating under the curve 
To be added

###Juvenile and Adult Ratio
```{r,message=F,warning=F}
ggplot(data, aes(x=sample_date, y=j_ratio, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + facet_grid(rows= vars(temp), cols = vars(community))+theme_bw()

#Look at j_ratio for just D vs DM buckets
ggplot(filter(summary,community=="D"|community=="DM"),aes(x=sample_date,y=mean_j_ratio,color=community))+ geom_point() + geom_line(aes(group=community)) + geom_errorbar(aes(ymin=mean_j_ratio-se_j_ratio,ymax=mean_j_ratio+se_j_ratio),width=0.2) + facet_wrap(.~temp,nrow=2)

ggplot(filter(summary,community=="D"),aes(x=sample_date,y=mean_j_ratio,color=temp))+ geom_point() + geom_line(aes(group=temp)) + geom_errorbar(aes(ymin=mean_j_ratio-se_j_ratio,ymax=mean_j_ratio+se_j_ratio),width=0.2)+theme_bw()
```

##YSI data

ysi
```{r}
ysi_summ <- ysi %>% group_by(bucket, Date) %>% summarize(temp_f = mean(temp_f),
                                                         temp_c = mean(temp_c),
                                                         do_per = mean(do_per),
                                                         do_conc = mean(do_conc),
                                                         ph = mean(ph),
                                                         ph_mv = mean(ph_mv),
                                                         chlor_rfu = mean(chlor_rfu),
                                                         chlor_conc = mean(chlor_conc))
ysi_summ %<>% left_join(.,meta, by = "bucket")
ggplot(ysi_summ,aes(x=chlor_conc,y=chlor_rfu))+geom_point()
```


temp
```{r}
# ggplot(ysi_summ, aes(x=Date, y=temp_c, color = bucket)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + theme(legend.position = "none")
ggplot(ysi_summ, aes(x=Date, y=temp_c, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))
```

```{r}
ysi_summ %>% filter(pool == 7) %>% ggplot(., aes(x=Date, y=temp_c, color = bucket)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + theme(legend.position = "none")
```

do_per
```{r}
# ggplot(ysi_summ, aes(x=Date, y=do_per, color = community)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_wrap(vars(temp))
#
# ggplot(ysi_summ, aes(x=Date, y=do_per, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_wrap(vars(community))
ggplot(ysi_summ, aes(x=Date, y=do_per, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))+theme_bw()
```

ph
```{r}
# ggplot(ysi_summ, aes(x=Date, y=ph, color = community)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_wrap(vars(temp))
#
# ggplot(ysi_summ, aes(x=Date, y=ph, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_wrap(vars(community))
ggplot(ysi_summ, aes(x=Date, y=ph, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))+theme_bw()

ggplot(filter(ysi_summ,bucket=="26"|bucket=="30"), aes(x=Date, y=ph, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))
```

chlor_conc
```{r}
# ggplot(ysi_summ, aes(x=Date, y=chlor_conc, color = community)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() +  facet_wrap(vars(temp))
#
# ggplot(ysi_summ, aes(x=Date, y=chlor_conc, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() +  facet_wrap(vars(community))
ggplot(ysi_summ, aes(x=Date, y=chlor_conc, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community)) + scale_y_continuous(trans='log10')+theme_bw()

#figure out later
#ggplot(filter(ysi_summ,bucket=="26"|bucket=="28",bucket=="30"), aes(x=Date, y=chlor_conc, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_wrap(.~bucket)

min_chlor<-min(ysi_summ$chlor_conc)
ysi_summ$log_chlor<-log10(ysi_summ$chlor_conc+min_chlor+0.0001)

```
according to temp
```{r}
ggplot(ysi_summ, aes(x=Date, y=temp_c, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=do_per, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=ph_mv, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=ph, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=chlor_conc, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=chlor_rfu, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(temp), cols = vars(community))+ scale_y_continuous(trans='log10')
```

according to pool
```{r}
ggplot(ysi_summ, aes(x=Date, y=temp_c, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(pool), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=do_per, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(pool), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=ph, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(pool), cols = vars(community))
ggplot(ysi_summ, aes(x=Date, y=chlor_conc, color = temp)) + geom_point() + geom_line(aes(group=bucket)) + scale_color_discrete() + facet_grid(rows=vars(pool), cols = vars(community))
```


##Looking at density vs prevalence:  

```{r}
#plotting each time point (sampling) as a separate point with total density vs density
data %>% filter(community!="CM") %>%
  filter(daphnia_prev>0) %>%
  mutate(bucket56 = ifelse(bucket=="56","56","not")) %>%
  ggplot(aes(x=daphnia,y=daphnia_prev,color=bucket56)) +
  geom_point() +
  theme_classic() +
  facet_wrap(~temp) +
  ylab("Infection prevalence") + xlab("Total Daphnia density")

#now with number of infected individuals
data %>% filter(community!="CM") %>%
  mutate(bucket56 = ifelse(bucket=="56","56","not")) %>%
  filter(daphnia_prev>0) %>%
  ggplot(aes(x=daphnia,y=daphnia_inf,color=bucket56)) +
  geom_point() +
  theme_classic() +
  facet_wrap(~temp) +
  ylab("Number of infected individuals") + xlab("Total Daphnia density")

data %>% filter(community!="CM") %>%
  #mutate(bucket56 = ifelse(bucket=="56","56","not")) %>%
  filter(daphnia_prev>0) %>%
  ggplot(aes(x=daphnia_inf,y=daphnia_prev)) +
  geom_point() +
  theme_classic() +
  facet_wrap(~temp,nrow=1) +
  ylab("Infection prevalence") + xlab("Number of infected Daphnia") +
  geom_smooth(method='lm',se=F,color="gray60")

```

